---
title: "hw2"
author: "andres matsin"
date: "23 September 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Exercise 1

32561 rows and 12 columns. All the columns are treated as character type. Must convert some to numeric values.
Got help from : https://stackoverflow.com/questions/22772279/converting-multiple-columns-from-character-to-numeric-format-in-r

```{r}
tabcols <- names(adult)
cols
cols.char <- c("workclass","education","occupation","native.country","male","female")
cols.char
cols.num <- cols[!cols %in% cols.char]
cols.num
adult.char <- adult[cols.char]
adult.num <- as.data.frame(lapply(adult[cols.num],as.numeric))
adult2 <- cbind(adult.char, adult.num)
adult <-adult2
```


Data was tab separated. Replaced then ? with NA. Cleaned the spaces just in case.

```{r}
adult <- read_delim("~/Data Mining/HW2/adult.csv", 
+     "\t", escape_double = FALSE, trim_ws = TRUE)

adult[adult == "?"] <- NA

adult[] <- lapply(adult, gsub, pattern = " ", replacement = "", fixed = TRUE)
```

*Age* column
Theres -57 value in the Age column and another with 320. Replaced them with NA. 2 rows affected
```{r}
adult$age[adult$age < 0] <- NA
adult$age[adult$age > 129] <- NA

```

*Worklass* column replaced privat with Private. 10 rows affected.

 adult[] <- lapply(adult, gsub, pattern = "privat", replacement = "Private", fixed = TRUE)
 
*native.country* renamed to country
```{r}
 names(adult)[names(adult) == 'native.country'] <- 'country'
 data$country[data$country=="Unitedstates"] = "United-States"
 adult$country[adult$country=="Unitedstates"] = "United-States"
```


# Exercise 2
a)
```{r}
 res <- adult %>%
          group_by(education, occupation) %>%
          summarise(avg_salary = mean(salaries, na.rm=TRUE)) %>%
      arrange(desc(avg_salary)) %>% spread(occupation, avg_salary)
```
b)

```{r}
 res <- res  %>% gather(colnames(res)[2:16], key="Occupation", value="average_salary")
```
c)
```{r}
ggplot(data=res) + geom_tile(aes(x=res$education,y=res$occupation,fill=res$average_salary))
```

d) 
```{r}
res$education <-  factor(res$education, levels=c("Preschool","1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th", "12th","Prof-school", "Some-college", "Bachelors", "HS-grad", "Masters", "Assoc-adm", "Assoc-voc", "Doctorate" ), ```
ordered=TRUE)
```
e)

```{r}
library(ggplot2)
ggplot(data = iris, aes(x = Sepal.Length, y = Sepal.Width, colour = Species)) + 
         geom_point() + 
         theme_bw()
```

# Exercise 3
Please provide comprehensive answers to all questions as we **actually** going to **read them all**.

# Exercise 4
You may talk to your peers but make sure to submit unique solutions, otherwise it would be considered as **plagiarism**.
