---
title: "hw2"
author: "andres matsin"
date: "23 September 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Exercise 1

Data was tab separated. Replaced then ? with NA. Cleaned the spaces just in case.

```{r}
library(readr)
adult <- read_delim("adult.csv", "\t", escape_double = FALSE, trim_ws = TRUE)

adult[adult == "?"] <- NA

adult[] <- lapply(adult, gsub, pattern = " ", replacement = "", fixed = TRUE)
```

32561 rows and 12 columns. All the columns are treated as character type. Must convert some to numeric values.
Got help from : https://stackoverflow.com/questions/22772279/converting-multiple-columns-from-character-to-numeric-format-in-r

```{r}
cols <- names(adult)
cols
cols.char <- c("workclass","education","occupation","native.country","male","female")
cols.char
cols.num <- cols[!cols %in% cols.char]
cols.num
adult.char <- adult[cols.char]
adult.num <- as.data.frame(lapply(adult[cols.num],as.numeric))
adult2 <- cbind(adult.char, adult.num)
adult <-adult2
```

*Age* column
Theres -57 value in the Age column and another with 320. Replaced them with NA. 2 rows affected
```{r}
adult$age[adult$age < 0] <- NA
adult$age[adult$age > 129] <- NA
```
Values: 17-90
RATIO
NA's   :99 

*Worklass* is nominal. column replaced privat with Private. 10 rows affected.
```{r}
 adult$workclass[adult$workclass=="privat"] = "Private"
```

```{r}
unique(adult[["workclass"]])
barplot(prop.table(table(adult[["workclass"]])))
```

Values 
*native.country* nominal, renamed to country, corrected United-States missspelling. 882 + 891 rows affected.
```{r}
 names(adult)[names(adult) == 'native.country'] <- 'country'
 adult$country[adult$country=="Unitedstates"] = "United-States"
adult$country[adult$country=="UnitedStates"] = "United-States"
```
*Education* ordinal.
```{r}
barplot(prop.table(table(adult[["education"]])))
```
*capital-gain* ratio
There seems to be a cut-off at 99999.
```{r}
library(ggplot2)
o<-subset(adult, adult$capital.gain > 0)
 o$capital.gain <- as.numeric(o$capital.gain)
 ggplot(o, aes(o$capital.gain)) + geom_histogram(binwidth=500)
```
*capital-loss* ratio

```{r}
o <- subset(adult, adult$capital.loss > 0)
o$capital.loss <- as.numeric(o$capital.loss)
ggplot(o, aes(o$capital.loss)) + geom_histogram(binwidth=100)
```
From 0-4000
*jobsatiscation* ratio
values 0-15
```{r}
hist(adult$jobsatisfaction)
```
# Exercise 2
a)

```{r}
library(tidyr)
library(dplyr)
 res <- adult %>%
          group_by(education, occupation) %>%
          summarise(avg_salary = mean(salaries, na.rm=TRUE)) %>%
      arrange(desc(avg_salary)) %>% spread(occupation, avg_salary)
```
b)

```{r}
 res <- res  %>% gather(colnames(res)[2:16], key="Occupation", value="average_salary")
```
c)
```{r}
ggplot(data=res) + geom_tile(aes(x=res$education,y=res$Occupation, fill=res$average_salary)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))  + labs(x="Education", colour="Average Salary", y="Occupation")
```

d) Not sure about the 
```{r}
res$education <- factor(res$education, levels=c("Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th", "12th", "HS-grad", "Prof-school", "Assoc-acdm", "Assoc-voc", "Some-college", "Bachelors", "Masters", "Doctorate"), ordered=TRUE)

ggplot(data=res) + geom_tile(aes(x=res$education,y=res$Occupation, fill=res$average_salary)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) + labs(x="Education", colour="Average Salary", y="Occupation")
```


Exercise 3
a) user_id is nominal, the id of the user
eval_set is nominal, evaluation set belonging. is data used to train, test or develop models
order_number is nominal 
order_dow is nominal time of the day. values from 0-6 though
order_hour_of_the_day is ordinal the hour of the day
days_since_last_order is interval - days since last order
product_id is nominal
add_to_car_order is ratio - in what order were the items added
reordered is nominal - binary. is the product reordered - has this product been ordered before
product_name is nominal 
aisle_id is nominal
department_id is nominal
aisle and department are both nominal

b)
```{r}
library(readr)
library(ggplot2)
instacart_100k <- read_csv("~/RProjects/rhw/instacart_100k.csv")
o <- instacart_100k

o$order_hour_of_day <- as.numeric(o$order_hour_of_day)
ggplot(o, aes(x=order_hour_of_day)) +
  geom_histogram(binwidth = 1) +
  facet_grid(. ~ order_dow)
```

I tried to guess what day is nr 0 and i am guessing its sunday, because it has the most activity at night, when people dont need to go to work in the morning.

```{r}
p <- subset(o, o$add_to_cart_order < 6)

ggplot(p, aes(x=add_to_cart_order)) +
   geom_histogram(binwidth = 1) +
   facet_grid(. ~ department)
```
A look at add to cart order (first five) by department. Some items are more likely to enter cart as the first item. 
